Proofs as Programs in Classical Logic\\Notes

### 2022-03-21 Notes/notes.tex
asdf
Grant Proposal First Draft

\usepackage{amsmath,amssymb,amsthm}
\usepackage{ wasysym }
\usepackage{ stmaryrd }
\usepackage{ mathpartir }
\usepackage{xcolor}
\usepackage{bussproofs}
\usepackage[left=2.00cm, right=2.00cm, top=2.00cm,bottom=2.0cm]{geometry}
\usepackage{pdflscape}
\usepackage{xspace}
\usepackage{todonotes}
\usepackage{stackengine}

@@

\begin{tabular}{lll}
	\AxiomC{\hphantom{x}}
	\RightLabel{Ax ($P$ atomic)}
	\UnaryInfC{$P,\Gamma\Rightarrow \Delta, P$}
	\DisplayProof&
	\AxiomC{\hphantom{x}}
	\RightLabel{L$\bot$}
	\UnaryInfC{$\bot,\Gamma\Rightarrow\Delta$}
	\DisplayProof&
\\&&\\
	\AxiomC{$A, B,\Gamma\Rightarrow\Delta$}
	\RightLabel{L$\wedge$}
	\UnaryInfC{$A\wedge B, \Gamma\Rightarrow \Delta$}
	\DisplayProof&
	\AxiomC{$\Gamma\Rightarrow\Delta, A$}
	\AxiomC{$\Gamma\Rightarrow\Delta, B$}
	\RightLabel{R$\wedge$}
	\BinaryInfC{$\Gamma\Rightarrow \Delta, A\wedge B$}
	\DisplayProof&
\\&&\\
	\AxiomC{$A, \Gamma\Rightarrow\Delta$}
	\AxiomC{$B, \Gamma\Rightarrow\Delta$}
	\RightLabel{L$\vee$}
	\BinaryInfC{$A\vee B, \Gamma\Rightarrow \Delta$}
	\DisplayProof&
	\AxiomC{$\Gamma\Rightarrow\Delta, A, B$}
	\RightLabel{R$\vee$}
	\UnaryInfC{$\Gamma\Rightarrow \Delta, A\vee B$}
	\DisplayProof&
\\&&\\
	\AxiomC{$A\to B, \Gamma\Rightarrow\Delta, A$}
	\AxiomC{$B, \Gamma\Rightarrow\Delta$}
	\RightLabel{L$\to$}
	\BinaryInfC{$A\to B, \Gamma\Rightarrow \Delta$}
	\DisplayProof&
	\AxiomC{$A,\Gamma\Rightarrow\Delta, B$}
	\RightLabel{R$\to$}
	\UnaryInfC{$\Gamma\Rightarrow \Delta, A\to B$}
	\DisplayProof&
\\&&\\
	\AxiomC{$\forall xA, A[t/x], \Gamma\Rightarrow\Delta$}
	\RightLabel{L$\forall$}
	\UnaryInfC{$\forall xA, \Gamma\Rightarrow \Delta$}
	\DisplayProof&
	\AxiomC{$\Gamma\Rightarrow\Delta, A[a/x]$}
	\RightLabel{R$\forall$}
	\UnaryInfC{$\Gamma\Rightarrow \Delta, \forall xA$}
	\DisplayProof&
\\&&\\
	\AxiomC{$A[a/x], \Gamma\Rightarrow\Delta$}
	\RightLabel{L$\exists$}
	\UnaryInfC{$\exists xA, \Gamma\Rightarrow \Delta$}
	\DisplayProof&
	\AxiomC{$\Gamma\Rightarrow\Delta, A[t/x], \exists xA$}
	\RightLabel{R$\exists$}
	\UnaryInfC{$\Gamma\Rightarrow \Delta, \exists xA$}
	\DisplayProof&
\\&&\\
	\AxiomC{$t = t, \Gamma\Rightarrow \Delta$}
	\RightLabel{Ref}
	\UnaryInfC{$\Gamma\Rightarrow \Delta$}
	\DisplayProof&
	\AxiomC{$s = t, P[s/x], P[t/x], \Gamma\Rightarrow \Delta$}
	\RightLabel{Ref}
	\UnaryInfC{$s = t, P[s/x],\Gamma\Rightarrow \Delta$}
	\DisplayProof&
\\&&\\
\end{tabular}

where in R$\forall$ and L$\exists$ $a$ is a free variable not occurring in the conclusion.

@@

	\hspace{-1cm}
	\AxiomC{}
	\RightLabel{Ax}
	\UnaryInfC{\scriptsize{\stackanchor{$P(a, b)\to Q(a, f(b)), \forall x\forall y(P(x, y)\to Q(x, f(y)))$}{$\forall x\exists yP(x, y), P(a, b), \forall y(P(a, y)\to Q(a, f(y)))$}$\Rightarrow$\stackanchor{$P(a, b), Q(a, f(b))$}{$\exists yQ(a, f(y))$}}}
	\RightLabel{Ax}
	\AxiomC{}
	\UnaryInfC{\scriptsize{\stackanchor{$Q(a, f(b)), P(a, b), \forall y(P(a, y)\to Q(a, f(y)))$}{$\forall x\exists yP(x, y), \forall x\forall y(P(x, y)\to Q(x, f(y)))$}$\Rightarrow$\stackanchor{$Q(a, f(b))$}{$\exists yQ(a, f(y))$}}}
	\RightLabel{L$\to$}
	\BinaryInfC{\stackanchor{$\forall y(P(a, y)\to Q(a, f(y))), P(a, b)\to Q(a, f(b))$}{$ \forall x\exists yP(x, y), P(a, b), \forall x\forall y(P(x, y)\to Q(x, f(y))) $}$\Rightarrow Q(a, f(b)), \exists yQ(a, f(y))$}
	\RightLabel{L$\forall$}
	\UnaryInfC{\stackanchor{$\forall x\forall y(P(x, y)\to Q(x, f(y))), P(a, b)$}{$\forall y(P(a, y)\to Q(a, f(y))), \forall x\exists yP(x, y)$}$\Rightarrow Q(a, f(b)), \exists yQ(a, f(y))$}
	\RightLabel{L$\forall$}
	\UnaryInfC{$P(a, b),\forall x\exists yP(x, y), \forall x\forall y(P(x, y)\to Q(x, f(y)))\Rightarrow Q(a, f(b)), \exists yQ(a, f(y))$}
	\RightLabel{R$\exists$}
	\UnaryInfC{$P(a, b),\forall x\exists yP(x, y), \forall x\forall y(P(x, y)\to Q(x, f(y)))\Rightarrow \exists yQ(a, f(y))$}
	\RightLabel{L$\exists$}
	\UnaryInfC{$\exists yP(a, y),\forall x\exists yP(x, y), \forall x\forall y(P(x, y)\to Q(x, f(y)))\Rightarrow \exists yQ(a, f(y))$}
	\RightLabel{L$\forall$}
	\UnaryInfC{$\forall x\exists yP(x, y),\forall x\forall y(P(x, y)\to Q(x, f(y)))\Rightarrow \exists yQ(a, f(y))$}
	\DisplayProof
\end{example}

@@

\begin{center}
	\AxiomC{$A, \Gamma\Rightarrow B$}
	\RightLabel{R$\to$}
	\UnaryInfC{$\Gamma\Rightarrow\Delta, A\to B$}
	\DisplayProof\hspace{2cm}
	\AxiomC{$\Gamma\Rightarrow A[a/x]$}
	\RightLabel{R$\forall$}
	\UnaryInfC{$\Gamma\Rightarrow\Delta, \forall xA$}
	\DisplayProof
\end{center}

@@

	$${\hspace*{-2.5cm}}\mathfrak f_3\left(\usebox\InnerProofga, \Gamma'\Rightarrow\Delta'\right)= 
	\AxiomC{$\mathfrak f_3\left(\usebox\InnerProofgb, \Gamma'\Rightarrow \Delta', A\right)$}\AxiomC{$\mathfrak f_3\left(\usebox\InnerProofgb, B, \Gamma'\setminus\{A\to B\}\Rightarrow \Delta'\right)$}\RightLabel{L$\to$}\BinaryInfC{$\Gamma'\Rightarrow \Delta'$}\DisplayProof$$

@@

$$\mathfrak{f_2}\left(\AxiomC{$\Rightarrow A\to B$}
\RightLabel{R$\to$}
\UnaryInfC{$A\Rightarrow B$}
\AxiomC{$A\to B\Rightarrow$}
\RightLabel{L$\to$1}
\UnaryInfC{$B\Rightarrow$}
\RightLabel{Res}
\BinaryInfC{A$\Rightarrow$}
\AxiomC{$A\to B\Rightarrow$}
\RightLabel{L$\to$1}
\UnaryInfC{$\Rightarrow A$}
\RightLabel{Res}
\BinaryInfC{$\Rightarrow$}
\DisplayProof\right)$$ can be, if we match on L$\to$ first, the intuitionistically valid $$\AxiomC{$A\to B, A\Rightarrow B, A$}\AxiomC{$A, B\Rightarrow B$}\RightLabel{R$\to$}\BinaryInfC{$A\to B, A\Rightarrow B$}\RightLabel{L$\to$}\UnaryInfC{$A\to B\Rightarrow A\to B$}\DisplayProof$$
or, if we match on R$\to$ first, the not intuitionistically valid $$\AxiomC{$A\to B, A\Rightarrow B, A$}\RightLabel{R$\to$}\UnaryInfC{$A\to B\Rightarrow A\to B, A$}\AxiomC{$A, B\Rightarrow B$}\RightLabel{R$\to$}\UnaryInfC{$B\Rightarrow A\to B$}\RightLabel{L$\to$}\BinaryInfC{$A\to B\Rightarrow A\to B$}\DisplayProof$$
a
\\

Now towards the main result:
\begin{definition}
	A formula is \textit{relevant} if it can constructed by the following steps:
	\begin{itemize}
		\item $\bot$ is a relevant
		\item If $B$ is relevant and $A$ is arbitrary $A\to B$ is relevant.
		\item if $A$ is relevant then $\forall x A[x/a]$ is relevant.
	\end{itemize}
	it is \textit{irrelevant} if it is not relevant.
\end{definition}

Intuitively a relevant formula is a formula that ends in $\bot$.

We can now by simultaneous induction define the notions of goal formulas and definite formulas

\begin{definition}
	A formula is a \textit{Goal Formula} if it can be constructed by the following steps:
	\begin{itemize}
		\item Any atomic formula is a goal formula
		\item For a definite formula $D$, goal formula $G$, if $D$ is relevant or quantifier-free, $D\to G$ is a goal formula.
		\item If $G$ is a goal formula then $\forall xG[x/a]$ is a goal formula.
	\end{itemize}
	A formula is a \textit{Definite Formula} if it can be constructed by the following steps:
	\begin{itemize}
		\item Any atomic formula is a definite formula
		\item For a definite formula $D$, goal formula $G$, if $D$ is relevant or $G$ irrelevant, $G\to D$ is a definite formula.
		\item If $D$ is a definite formula then $\forall xD[x/a]$ is a definite formula.
	\end{itemize}
\end{definition}

\begin{theorem}
	Let $D_1,\dots, D_n$ be definite formulas and $G$ a goal formula. If $D_1,\dots,D_n\Rightarrow G$ has a resolution proof then $D_1,\dots,D_n\Rightarrow G$ is provable in \Gi.
\end{theorem}

\begin{lemma}
	Let $A_1,\dots, A_n\Rightarrow B_1, \dots, B_m$ be the label in the node of a resolution refutation of $\{\Rightarrow D_i\}\cup\{G\}$. Then each $A_i$ is a goal forumla and each $B_i$ is definite.
\end{lemma}
\begin{proof}
	Induction.
\end{proof}
\pagebreak

>>>>

\usepackage{amsmath,amssymb,amsthm}
\usepackage{ wasysym }
\usepackage{ stmaryrd }
\usepackage{ mathpartir }
\usepackage{xcolor}
\usepackage{bussproofs}
\usepackage[left=2.00cm, right=2.00cm, top=2.00cm,bottom=2.0cm]{geometry}
\usepackage{pdflscape}
\usepackage{xspace}
\usepackage{todonotes}
\usepackage{stackengine}
\usepackage{tikz}
\usepackage{tikz-qtree}

@@

\begin{tabular}{lll}
	\AxiomC{\hphantom{x}}
	\RightLabel{Ax ($P$ atomic)}
	\UnaryInfC{$P,\Gamma\Rightarrow \Delta, P$}
	\DisplayProof&
	\AxiomC{\hphantom{x}}
	\RightLabel{L$\bot$}
	\UnaryInfC{$\bot,\Gamma\Rightarrow\Delta$}
	\DisplayProof&
\\&&\\
	\AxiomC{$A, B,\Gamma\Rightarrow\Delta$}
	\RightLabel{L$\wedge$}
	\UnaryInfC{$A\wedge B, \Gamma\Rightarrow \Delta$}
	\DisplayProof&
	\AxiomC{$\Gamma\Rightarrow\Delta, A$}
	\AxiomC{$\Gamma\Rightarrow\Delta, B$}
	\RightLabel{R$\wedge$}
	\BinaryInfC{$\Gamma\Rightarrow \Delta, A\wedge B$}
	\DisplayProof&
\\&&\\
	\AxiomC{$A, \Gamma\Rightarrow\Delta$}
	\AxiomC{$B, \Gamma\Rightarrow\Delta$}
	\RightLabel{L$\vee$}
	\BinaryInfC{$A\vee B, \Gamma\Rightarrow \Delta$}
	\DisplayProof&
	\AxiomC{$\Gamma\Rightarrow\Delta, A, B$}
	\RightLabel{R$\vee$}
	\UnaryInfC{$\Gamma\Rightarrow \Delta, A\vee B$}
	\DisplayProof&
\\&&\\
	\AxiomC{$\Gamma\Rightarrow\Delta, A$}
	\AxiomC{$B, \Gamma\Rightarrow\Delta$}
	\RightLabel{L$\to$}
	\BinaryInfC{$A\to B, \Gamma\Rightarrow \Delta$}
	\DisplayProof&
	\AxiomC{$A,\Gamma\Rightarrow\Delta, B$}
	\RightLabel{R$\to$}
	\UnaryInfC{$\Gamma\Rightarrow \Delta, A\to B$}
	\DisplayProof&
\\&&\\
	\AxiomC{$A[t/x], \Gamma\Rightarrow\Delta$}
	\RightLabel{L$\forall$}
	\UnaryInfC{$\forall xA, \Gamma\Rightarrow \Delta$}
	\DisplayProof&
	\AxiomC{$\Gamma\Rightarrow\Delta, A[a/x]$}
	\RightLabel{R$\forall$}
	\UnaryInfC{$\Gamma\Rightarrow \Delta, \forall xA$}
	\DisplayProof&
\\&&\\
	\AxiomC{$A[a/x], \Gamma\Rightarrow\Delta$}
	\RightLabel{L$\exists$}
	\UnaryInfC{$\exists xA, \Gamma\Rightarrow \Delta$}
	\DisplayProof&
	\AxiomC{$\Gamma\Rightarrow\Delta, A[t/x]$}
	\RightLabel{R$\exists$}
	\UnaryInfC{$\Gamma\Rightarrow \Delta, \exists xA$}
	\DisplayProof&
\\&&\\
	\AxiomC{$t = t, \Gamma\Rightarrow \Delta$}
	\RightLabel{Ref}
	\UnaryInfC{$\Gamma\Rightarrow \Delta$}
	\DisplayProof&
	\AxiomC{$s = t, P[s/x], P[t/x], \Gamma\Rightarrow \Delta$}
	\RightLabel{Ref}
	\UnaryInfC{$s = t, P[s/x],\Gamma\Rightarrow \Delta$}
	\DisplayProof&
\\&&\\
\end{tabular}

where in R$\forall$ and L$\exists$ $a$ is a free variable not occurring in the sequent otherwise.

@@

	\hspace{-.5cm}
	\AxiomC{}
	\RightLabel{Ax}
	\UnaryInfC{\scriptsize{\stackanchor{$\forall x\forall y(P(x, y)\to Q(x, f(y))), P(a, b)$}{$\forall x\exists yP(x, y), \forall y(P(a, y)\to Q(a, f(y)))$}$\Rightarrow$\stackanchor{$P(a, b), Q(a, f(b))$}{$\exists yQ(a, f(y))$}}}
	\RightLabel{Ax}
	\AxiomC{}
	\UnaryInfC{\scriptsize{\stackanchor{$Q(a, f(b)), P(a, b), \forall y(P(a, y)\to Q(a, f(y)))$}{$\forall x\exists yP(x, y), \forall x\forall y(P(x, y)\to Q(x, f(y)))$}$\Rightarrow$\stackanchor{$Q(a, f(b))$}{$\exists yQ(a, f(y))$}}}
	\RightLabel{L$\to$}
	\BinaryInfC{\stackanchor{$\forall y(P(a, y)\to Q(a, f(y))), P(a, b)\to Q(a, f(b))$}{$ \forall x\exists yP(x, y), P(a, b), \forall x\forall y(P(x, y)\to Q(x, f(y))) $}$\Rightarrow Q(a, f(b)), \exists yQ(a, f(y))$}
	\RightLabel{L$\forall$}
	\UnaryInfC{\stackanchor{$\forall x\forall y(P(x, y)\to Q(x, f(y))), P(a, b)$}{$\forall y(P(a, y)\to Q(a, f(y))), \forall x\exists yP(x, y)$}$\Rightarrow Q(a, f(b)), \exists yQ(a, f(y))$}
	\RightLabel{L$\forall$}
	\UnaryInfC{$P(a, b),\forall x\exists yP(x, y), \forall x\forall y(P(x, y)\to Q(x, f(y)))\Rightarrow Q(a, f(b)), \exists yQ(a, f(y))$}
	\RightLabel{R$\exists$}
	\UnaryInfC{$P(a, b),\forall x\exists yP(x, y), \forall x\forall y(P(x, y)\to Q(x, f(y)))\Rightarrow \exists yQ(a, f(y))$}
	\RightLabel{L$\exists$}
	\UnaryInfC{$\exists yP(a, y),\forall x\exists yP(x, y), \forall x\forall y(P(x, y)\to Q(x, f(y)))\Rightarrow \exists yQ(a, f(y))$}
	\RightLabel{L$\forall$}
	\UnaryInfC{$\forall x\exists yP(x, y),\forall x\forall y(P(x, y)\to Q(x, f(y)))\Rightarrow \exists yQ(a, f(y))$}
	\DisplayProof
\end{example}

@@

\begin{center}
	\AxiomC{$A, \Gamma\Rightarrow B$}
	\RightLabel{R$\to_1$}
	\UnaryInfC{$\Gamma\Rightarrow\Delta, A\to B$}
	\DisplayProof\hspace{2cm}
	\AxiomC{$\Gamma\Rightarrow A[a/x]$}
	\RightLabel{R$\forall_1$}
	\UnaryInfC{$\Gamma\Rightarrow\Delta, \forall xA$}
	\DisplayProof
\end{center}

@@

	$${\hspace*{-2.95cm}}\mathfrak f_3\left(\usebox\InnerProofga, \Gamma'\Rightarrow\Delta'\right)= 
	\AxiomC{$\mathfrak f_3\left(\usebox\InnerProofgb, \Gamma'\setminus\{A\to B\}\Rightarrow \Delta', A\right)$}\AxiomC{$\mathfrak f_3\left(\usebox\InnerProofgb, B, \Gamma'\setminus\{A\to B\}\Rightarrow \Delta'\right)$}\RightLabel{L$\to$}\BinaryInfC{$\Gamma'\Rightarrow \Delta'$}\DisplayProof$$

@@

$$\mathfrak{f_2}\left(\AxiomC{$\Rightarrow A\to B$}
\RightLabel{R$\to$}
\UnaryInfC{$A\Rightarrow B$}
\AxiomC{$A\to B\Rightarrow$}
\RightLabel{L$\to$1}
\UnaryInfC{$B\Rightarrow$}
\RightLabel{Res}
\BinaryInfC{A$\Rightarrow$}
\AxiomC{$A\to B\Rightarrow$}
\RightLabel{L$\to$1}
\UnaryInfC{$\Rightarrow A$}
\RightLabel{Res}
\BinaryInfC{$\Rightarrow$}
\DisplayProof\right)$$ can be, if we match on L$\to$ first, the intuitionistically valid $$\AxiomC{$A\to B, A\Rightarrow B, A$}\AxiomC{$A, B\Rightarrow B$}\RightLabel{R$\to$}\BinaryInfC{$A\to B, A\Rightarrow B$}\RightLabel{L$\to$}\UnaryInfC{$A\to B\Rightarrow A\to B$}\DisplayProof$$
or, if we match on R$\to$ first, the not intuitionistically valid $$\AxiomC{$A\to B, A\Rightarrow B, A$}\RightLabel{R$\to$}\UnaryInfC{$A\to B\Rightarrow A\to B, A$}\AxiomC{$A, B\Rightarrow B$}\RightLabel{R$\to$}\UnaryInfC{$B\Rightarrow A\to B$}\RightLabel{L$\to$}\BinaryInfC{$A\to B\Rightarrow A\to B$}\DisplayProof$$
\\

Now we may ask what structural conditions we have to impose on a sequent so that we can always extract an intuitionistic proof, i.e. every instance of R$\to$ and R$\to$ is a valid instance of R$\to_1$ and R$\forall_1$.

As usual we can associate with every sequent and formula a tree:

\begin{definition}
	A \textit{Formula Tree} is a labelled ordered rooted tree in which every node is labelled with a connector, quantifier or atom. We can define a function $\iota$ that transforms sequents and formulas into formula trees as follows:
	\begin{itemize}
		\item $\iota(A_1,\dots, A_n\Rightarrow B) = \iota(A_1,\dots, A_{n-1}\Rightarrow A_n\to B)$, $\iota(\Rightarrow B) = \iota(B)$
		\item $\iota(A\square B) = $\tikz[scale=0.9, baseline=-5mm, thick]{\Tree [.$\square$ [.$\iota(A)$  ] [.$\iota(B)$  ] ]} for $\square\in\{\wedge,\vee,\to\}$.
		\item $\iota(\square xA) = $\tikz[scale=0.9, baseline=-5mm, thick]{\Tree [.$\square x$ [.$\iota(A)$  ]]} for $\square\in\{\forall, \exists\}$.
		\item $\iota(A) = A$ for $A$ atomic.
	\end{itemize}
\end{definition}

\begin{definition}
	To each node of a formula tree we assign a \textit{Polarity} as follows:
	The polarity of the root node is $1$. If the label of a node with polarity $p$ is in $\{\wedge,\vee,\forall,\exists\}$ then the polarity of its children is also $p$. If the label of a node with polarity $p$ is $\to$ then then polarity of its left child is $-p$ and of its right child $p$.
\end{definition}
 We annotate the polarity of a node with a superscript $+$ or $-$.

\begin{example}
	For our running example $\forall x\exists yP(x, y), \forall x\forall y(P(x, y)\to Q(x, f(y)))\Rightarrow\forall x\exists Q(x, y)$ the associated formula tree is
	\begin{center}
		\begin{tikzpicture}
			\Tree [.$\to^+$ [.$\to^-$  [.$\forall x^+$ [.$\exists y^+$ $P(x,y)^+$ ] ] [.$\exists x^-$ [.$\forall y^-$ [.$\to^-$ $P(x,y)^+$ $Q(x,f(x))^-$ ] ] ] ]  [.$\forall x^+$  [.$\exists y^+$ $Q(x,y)^+$ ] ] ]
		\end{tikzpicture}
	\end{center}
\end{example}

We can now give sufficient conditions for when an extracted proof will always be intuitionstic:

\begin{theorem}
	Suppose in the formula tree of $\varphi$ there is no occurrence of $\to^+$ or $\forall^+$ below $\to^-$ or $\vee^+$. Then for any resolution proof $\mathcal{T}$ of $\varphi$ the extracted \Gc proof $\mathfrak{f}_2(\mathcal{T})$ is a valid \Gi proof.\\
\end{theorem}

This is however quite restrictive. Even quite simple, intuitionistically valid sequents like $$(A\to A)\to A\Rightarrow A$$ or $$(A\to B)\to C, A\to B\Rightarrow C$$ are not covered by the above theorem.\\
However by examining alternatives to rules R$\to_1$, R$\forall_1$ we can find more broadly applicable results.

\subsection{$\Pi_2$ sentences in PA}

\subsection{A refined algorithm}

Consider the calculus \Gi' obtained from \Gi by replacing R$\to_1$ and R$\forall_1$ with the rules 
\begin{center}
	\AxiomC{$A, \Gamma\Rightarrow \Delta, B$}
	\RightLabel{R$\to_2$}
	\UnaryInfC{$\Gamma\Rightarrow\Delta, A\to B$}
	\DisplayProof\hspace{2cm}
	\AxiomC{$\Gamma\Rightarrow \Delta,A[a/x]$}
	\RightLabel{R$\forall_2$}
	\UnaryInfC{$\Gamma\Rightarrow\Delta, \forall xA$}
	\DisplayProof
\end{center}
where the introductory axiom in the branch is not between atoms from $\Delta$ and $A$.

In we analyse our reasoning for theorem 5.7 under this calculus we identify for each formula precisely the atoms we may not resolve with each other.

Intuitively this corresponds to the fact that doubly negated atoms cannot resolve request.

We now propose a translation procedure for formulas $\varphi$ such that $\varphi^\circ$ is provable by resolution iff $\varphi$ is valid in \Gi'. The key insight is that atoms that occur left below $\to^-$ may not be resolved with atoms that occur below only positive nodes and atoms that occur below an initial $\vee^+$ may not be resolved with atoms from the other branch of $\vee^+$:

First rename every atom $P$ to $P^0$. Rename all atoms that do not occur below a $\to^-$ to $P^1$ and all positive atoms that occurr below only positive nodes to $P^{-1}$, these atoms are not affected by any further relabelling. If a node is labelled $\vee^+$ and every node above is positive relabel all atoms $P^i$ in the left branch to $P^{i+1}$ and in the right branch to $P^{i+2}$. The thus obtained formula we call $\varphi^\circ$ and for each atom $P$ we denote the maximum $i$ such that $P^i$ occurs in $\varphi^\circ$ with $\chi_P(\varphi)$.
\begin{example}The sequent $(A\to A)\to A\Rightarrow A$ gets transformed to $(A^0\to A^0)\to A^1\Rightarrow A^1$, the sequent $(A\to \bot)\to \bot\Rightarrow A$ gets transformed to $(A^0\to \bot)\to\bot\Rightarrow A^1$.

\begin{theorem}
	$\{P^i\to P^0\:|\:P \text{ atom in $\varphi$}, i\leq \chi_{P}(\varphi)\}\Rightarrow \varphi^\circ$ is provable by resolution iff $\varphi$ provable in \Gi'.
\end{theorem}

\end{example}
